---
title: "Practical Machine Learning Project"
author: "Felix E. Rivera-Mariani, PhD"
date: "February 25, 2016"
output: html_document
---
###Summary
The purpose of this project is to determine how well 6 participants perform a particular activity based on accelometers data on belt, forearm, arm, and dumbell. This will be accomplished in this report by a random forest classifier algorithm. 

The github repository is found at the following link: https://github.com/friveramariani/practica-machine-learning-prjt.git

The Caret package for subsetting the data, prepare a training subset, and cross validate the model. 

```{r, cache=TRUE}
library(caret)
library(randomForest)

# Global training data downloaded training data
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "./prjt-pml-training.csv")

data.training <- read.csv("./prjt-pml-training.csv")

na <- apply(data.training, 2, function(x) sum(x %in% c(NA, "")))
na
```

At this point, there are 152 variables because 8 "house keeping" variables were removed. Nevertheless, **there seems that some of the variables may contain a few data points (i.e. data sparsness).** These will be removed to avoid affecting the predictive value of the model. 
```{r, cache=TRUE}
index <- which(na == 0)

data.training <- data.training[,index]

data.training <- data.training[,8:60]
```

**The random forest algorithim** will be used because of the following:

1) Can handle large number of inputs
2) Provide a cross-validation component 
3) Can handle variables both categorial and unscalled
4) Can be used to estimate variable importance
5) Provides a probability output

A portion of the training subset (10%) will be used to determine the variables of importance. 
```{r, cache=TRUE}
model <- randomForest(classe~., data = data.training)
pred <- predict(model, data.training)
confusionMatrix(data.training$classe, pred)
```


```{r}
# Apply model to a different data subset
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "./prjt-pml-testing.csv")

library(randomForest)
data.testing <- read.csv("./prjt-pml-testing.csv")
data.testing <- data.testing[,index]
data.testing<- data.testing[,8:59]
data.testing$classe <- factor(nrow(data.testing))
levels(data.testing$classe) <- levels(data.testing$classe)
Test2 <- rbind(data.training[1,], data.testing) 
Test2 <- Test2[2:21,]
```

```{r}
TestModel <- predict(model, Test2)
TestModel
```